---
title: "PCA Website | Danny Pan"
output:
  html_document: default
---

## PCA

```{r, include = FALSE}
library(tidyverse)
```

```{r}
pca_compression <- function(X, k) {
  V = t(X) %*% X
  U = eigen(V)$vectors #eigenvectors
  lambda = eigen(V)$values #eigenvalues
  Z = X %*% U #pc scores
  Z_k = Z[, 1:k, drop = FALSE] #first k columns of pc scores
  U_k = U[, 1:k, drop = FALSE] #first k columns of eigenvectors
  X_k = Z_k %*% t(U_k) #rank k approx
  error = norm(X - X_k, type = "F")
  return(list(X_k = X_k, error = error,U1=U_k[,1],Z1=Z_k[,1]))
}


```

```{r echo = FALSE}
folder = list.files(path = "hw2_imgs/", pattern = "\\.csv$", full.names = TRUE)
for (file in folder) {
  X = as.matrix(read.csv(file), header = TRUE)
  # Original Image:
  image(X, col = grey(seq(0, 1, length = 256)))
  
  errors = numeric(10)
  
  for (k in 1:10) {
    output = pca_compression(X,k)
    image(output$X_k, col = grey(seq(0, 1, length = 256)), main = paste("k =", k))
    errors[k] <- output$error
  }
  
  errors_df <- tibble(k = 1:length(errors), error = errors)
  U1_df <- tibble(index = 1:length(output$U1), value = output$U1)
  Z1_df <- tibble(index = 1:length(output$Z1), value = output$Z1)
  
  print(ggplot(errors_df, aes(x = k, y = error)) + geom_line() + labs(title = "Error vs k", x = "k", y = "Frobenius Norm Error"))
  
  print(ggplot(U1_df, aes(x = index, y = value)) + geom_line() +
          labs(title = "First Eigenvector (u1)", x = "index", y = "Value"))
  
  print(ggplot(Z1_df, aes(x = index, y = value)) + geom_line() +
          labs(title = "First PC Score (z1)", x = "index", y = "Value"))
}
```

# 1st Eigenvector/PC Score Interpretations

Image 1: Noisy PC plot, no meaningful relationship within the image beyond \~row 78 as shown in eigenvector plot.

Image 2: Noisy PC plot, but eigenvector plot shows a mathematical trend allowing for compression.

Image 3: Smooth plots, signals structure within the image

Image 4: Plots signal slightly less structure compared to image 3 (puppy) but more contrast compared to images 1 & 2.

# Appropriate Values of K for Approximation

Plot 1: k = 1

Plot 2: k = 1

Plot 3: k = 25

Plot 4: k = 15
